---
title: "San Francisco Trees"
author: "Dean Marchiori"
date: "03/02/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The data this week comes from San Francisco's open data portal.  

```{r}
library(tidyverse)
library(sf)
library(mapview)
library(osmdata)
library(dodgr)
library(showtext)
```

Load data from https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-01-28/readme.md

```{r}
sf_trees <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-28/sf_trees.csv')
```

Filter out the trees with unknown locations and turn into a basic sf object  

```{r}
tree_geo <- sf_trees %>% 
  filter(!is.na(latitude) | !is.na(longitude)) %>% 
  st_as_sf(coords = c('longitude', 'latitude')) %>% 
  st_set_crs(4326) %>% 
  select(tree_id)
```

Start with the default guess for a bounding box for SFO. Manually tweaked
after inspecting with `mapview()`. 

Then get streets from osm.  

```{r}
# get rough auto bounding box
estimate_box <- osmdata::getbb("San Francisco, CA")

# tweak bounding box
estimate_box[1,1] <- -122.517
estimate_box[1,2] <- -122.350
estimate_box[2,1] <- 37.700
estimate_box[2,2] <- 37.815

# get streets
streets <- dodgr_streetnet(estimate_box)

street_geo <- streets %>% 
  select(osm_id, name)
```

Once we have trees and strees, we get the nearest street for each tree, then
summarise to count how many trees per street, then join back to the street
dataset for plotting.  

```{r}
trees_per_street <- tree_geo %>% 
  st_join(street_geo, join = st_nearest_feature) %>% 
  st_drop_geometry() %>% 
  count(osm_id)

  streets_w_counts <- street_geo %>% 
  left_join(trees_per_street) %>% 
  replace_na(list(n = 0))
```

make plot

```{r}
ggplot(streets_w_counts) +
  geom_sf(aes(alpha = n, colour = n), show.legend = FALSE) +
  guides(colour = "colourbar") + 
  scale_colour_viridis_c() +
  theme_minimal() +
  labs(title = "The Trees on San Francisco Streets",
       x = "",
       y = "",
       caption = "@deanmarchiori | data: data.sfgov.org") +
  scale_alpha_continuous(range = c(0.2, 0.8)) +
  annotate(geom = 'text', 
           x = -122.355, 
           y = 37.78, 
           label = "
             Passed in November 2016, Prop. E transferred the responsibility   
             of San Francisco’s nearly 125,000 trees from local residents back to the city.   
             San Francisco had unburdened its tree duty onto city dwellers back in 2011,   
             when it was facing a $380 million general fund deficit. But the decision by   
             Big Brother to dump Big Trees on San Franciscans did not work out,   
             to say the least.   
             
             The most common tree in San Francisco, with a population of 8,620,   
             is the London plane tree (Platanus acerifolia), a hybrid sycamore.  
             One of the world’s great street trees, it lines the Champs Élysées in Paris,   
             it’s found in a van Gogh painting, and it has been planted in cities 
             from Australia to Alaska.  
           
             - SF Weekly", 
           size = 3, 
           family = "Oxygen",
           colour = "#E2E2E2") +
  theme(
    title = element_text(family = "Oxygen", size = 24, color = '#E2E2E2'), 
    text = element_text(family = "Oxygen"),
    plot.background = element_rect(fill = "#5C5B57"),
    panel.background = element_rect(fill = "#5C5B57",  color = "#5C5B57"),
    panel.grid = element_blank(), 
    axis.ticks = element_blank(), 
    panel.border = element_blank(), 
    plot.caption = element_text(family = "Oxygen", size = 16, color = '#E2E2E2'))

ggsave(filename = 'img/sf-trees.png', width = 342, device = 'png', height = 374, units = "mm")
```

